---
title: "p8105_hw1_zc2959"
author: "Zhenghong Chen zc2959"
date: "2026-09-26"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```
## Penguins dataset description

The `penguins` dataset contains information on three penguin species: `r paste(levels(pull(penguins, species)), collapse = ", ")`. The penguins were observed on `r paste(levels(pull(penguins, island)), collapse = ", ")` islands. Important variables include `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`.

The dataset has `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)` mm.

## Penguins scatterplot

The following scatterplot shows `flipper_length_mm` vs `bill_length_mm`, with points colored by `species`.

```{r penguins_scatterplot}
penguins_scatterplot = ggplot(
  penguins,
  aes(
    x = bill_length_mm,
    y = flipper_length_mm,
    color = species
  )
) + 
  geom_point(na.rm = TRUE)
penguins_scatterplot
```

```{r save_penguins_scatterplot}
ggsave(
  "penguins_scatterplot.pdf",
  plot = penguins_scatterplot,
  height = 4,
  width = 6
)
```



# Problem 2

## Dataframe

```{r problem2_dataframe}
set.seed(1)

problem2_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c(
  "a", "b", "c", "d", "e",
  "f", "g", "h", "i", "j"
),
  vec_factor = factor(c(
    "A", "B", "C", "A", "B",
    "C", "A", "B", "C", "A"
  ))
)

```

## Variable means

```{r variable_means}
mean(pull(problem2_df, norm_samp))
mean(pull(problem2_df, norm_samp_pos))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, vec_factor))

```
### Description for variable means

The mean of `norm_samp` can be calculated because it is a numeric variable.

For `norm_samp_pos`, R treats TRUE as 1 and FALSE as 0, so

$$
\text{mean} = \frac{\#\{\text{TRUE}\}}{10},
$$

which is the proportion of observations for which `norm_samp > 0` is TRUE.

For `vec_char` and `vec_factor`, the mean is not defined directly because these variables are categorical rather than numeric. Therefore, R returns NA with a warning when attempting to calculate their means.


## Numeric conversion

```{r numeric_conversion, results = "hide", warning = FALSE}
as.numeric(pull(problem2_df, norm_samp_pos))
as.numeric(pull(problem2_df, vec_char))
as.numeric(pull(problem2_df, vec_factor))
```
### Description for numeric conversion

For `norm_samp_pos`, converting the logical values to numeric gives

$$
\text{TRUE} \rightarrow 1,
\qquad
\text{FALSE} \rightarrow 0.
$$

This explains why the mean of `norm_samp_pos` can be calculated.

For `vec_char`, converting to numeric produces NA values because the character strings do not represent numbers.

For `vec_factor`, converting to numeric produces integer codes for the factor levels. These values are category codes rather than meaningful numerical measurements.

Therefore, coercion helps explain why the mean can be calculated for a logical variable, but not directly for character or factor variables.



