---
title: "p8105_hw1_zc2959"
author: "Zhenghong Chen zc2959"
date: "2026-09-26"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```
## Penguins dataset description

The `penguins` dataset contains information on three penguin species: `r paste(levels(pull(penguins, species)), collapse = ", ")`. The penguins were observed on `r paste(levels(pull(penguins, island)), collapse = ", ")` islands. Important variables include `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`.

The dataset has `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)` mm.

## Penguins scatterplot

The following scatterplot shows `flipper_length_mm` vs `bill_length_mm`, with points colored by `species`.

```{r penguins_scatterplot}
penguins_scatterplot = ggplot(
  penguins,
  aes(
    x = bill_length_mm,
    y = flipper_length_mm,
    color = species
  )
) + 
  geom_point(na.rm = TRUE)
penguins_scatterplot
```

```{r save_penguins_scatterplot}
ggsave(
  "penguins_scatterplot.pdf",
  plot = penguins_scatterplot,
  height = 4,
  width = 6
)
```



# Problem 2

## Dataframe

```{r problem2_dataframe}
set.seed(1)

problem2_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c(
  "a", "b", "c", "d", "e",
  "f", "g", "h", "i", "j"
),
  vec_factor = factor(c(
    "A", "B", "C", "A", "B",
    "C", "A", "B", "C", "A"
  ))
)

```

## Variable means

```{r variable_means}
mean(pull(problem2_df, norm_samp))
mean(pull(problem2_df, norm_samp_pos))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, vec_factor))

```
### Description for variable means

The mean can be calculated for `norm_samp` and `norm_samp_pos`.

For the logical variable `norm_samp_pos`, R treats TRUE as 1 and FALSE as 0; therefore, its mean represents the proportion of TRUE values.

For `vec_char` and `vec_factor`, `mean()` returns NA and issues a warning because these variables are neither numeric nor logical.


## Numeric conversion

```{r numeric_conversion, results = "hide", warning = FALSE}
as.numeric(pull(problem2_df, norm_samp_pos))
as.numeric(pull(problem2_df, vec_char))
as.numeric(pull(problem2_df, vec_factor))
```
### Description for numeric conversion

When `norm_samp_pos` is converted to a numeric type, TRUE is converted to 1 and FALSE is converted to 0, which explains why its mean can be calculated.

Converting `vec_char` to a numeric type results in NA values because character strings do not represent numbers.

When converting `vec_factor` to a numeric type, integer codes corresponding to the factor levels are generated; these codes are internal category codes, not meaningful numerical measurements.

This helps explain why the mean of a logical variable can be calculated, whereas the mean of a character or factor variable cannot be calculated directly.




