---
title: "p8105_hw1_zc2959"
author: "Zhenghong Chen zc2959"
date: "2026-09-26"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```
## penguins dataset description

The `penguins` dataset contains information on three penguin species:`r paste(levels(pull(penguins,species)), collapse = ", ")`. The penguins were observed on `r paste(levels(pull(penguins,island)), collapse = ", ")` islands. Important variables include `species`, `island`, `bill_length_mm`,`bill_depth_mm`,`flipper_length_mm`, `body_mass_g`, `sex`, and `year`.

The dataset has `r nrow(penguins)` observations and `r ncol(penguins)` variables. The mean flipper length is `r round(mean(pull(penguins,flipper_length_mm), na.rm = TRUE),2)`mm.

## penguins scatterplot

```{r penguins_scatterplot}
penguins_scatterplot = ggplot(
  penguins,
  aes(
    x = bill_length_mm,
    y = flipper_length_mm,
    color = species
  )
) + 
  geom_point(na.rm = TRUE)
penguins_scatterplot
```

```{r save_penguins_scatterplot}
ggsave(
  "penguins_scatterplot.pdf",
  plot = penguins_scatterplot,
  height = 5,
  width = 7
)
```



# Problem 2